---
title: "Shopping Behavior - Data Visualization"
author: "Paria Ighanian"
date: "2023-11-27"
output: html_document

---

```{r}
library(ggplot2)
library(dplyr)
```


```{r}
data <- read.csv('shopping_behavior.csv')
dim(data)
```
The dataset consists of 3900 rows and 18 columns.

```{r}
str(data)

```
```{r}
colnames(data) <- c('customer_id', 'age', 'gender', 'item_purchased', 'category',
                  'purchase_amount_usd', 'location', 'size', 'color', 'season',
                  'review_rating', 'subscription_status', 'shipping_type',
                  'discount_applied', 'promo_code_used', 'previous_purchases',
                  'payment_method', 'frequency_of_purchases')

```



# Exploring Data
## Customer Distribution by Gender
We would like to understand the distribution of customers based on their gender. 

```{r}
df_percentage <- (data |>
  group_by(gender) |>
  summarise(percentage = n() / nrow(data) * 100))

ggplot(df_percentage, aes(x = gender, y=percentage, fill = gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_stack(vjust = 0.5)) +

  labs(title = "Customer Distribution by Gender",
       x = "Gender",
       y = "Percentage")+
  scale_y_continuous(labels = scales::percent_format(scale = 1))+
  theme(axis.title.y = element_text(margin = margin(r = 10)), legend.position = "none")


```

## Category Distribution by Gender
We would like to know if there are potential patterns among different age groups within each gender.
```{r}
percentage_data <- (data |>
  group_by(gender, category) |>
  summarise(count = n()) |>
  group_by(gender) |>
  mutate(percentage = (count / sum(count)) * 100))

ggplot(percentage_data, aes(x = gender, y = percentage, fill = category, label=sprintf("%.1f%%", percentage))) +
  geom_bar(position = "stack", stat = "identity", alpha = 0.7) +
    geom_text(position = position_stack(vjust = 0.5)) +

  labs(title = "Category Distribution by Gender",
       x = "Gender",
       y = "Percentage",
      fill = "Category") +
  scale_y_continuous(labels = scales::percent_format(scale = 1))+
  theme_minimal()
```

## Number of Purchases by Age
Delve into our customer demographics to discover the standout performers among all age groups.
```{r}
purchase_counts <- data |>
  group_by(age) |>
  summarise(purchase_count = n())

ggplot(purchase_counts, aes(x = factor(age), y = purchase_count)) +
  geom_col(fill = "lightblue") +
  geom_line(aes(group = 1), color = "blue", size = 1) +
  labs(title = "Number of Purchases by Age",
       x = "Age",
       y = "Number of Purchases") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


